---
title: '**CSC487**: Data Mining - Homework #2'
output:
  pdf_document:
    latex_engine: pdflatex
    extra_dependencies:
    - xcolor
    - hyperref
  word_document: default
  html_document:
    df_print: paged
fontsize: 12pt
graphics: yes
---

```{=latex}
\newcommand{\XB}{\color{black}}
\newcommand{\XBB}{\color{blue}}
\newcommand{\XV}{\color{violet}}
\newcommand{\XR}{\color{red}}

\begin{center}
  \XV\textit{\large{\href{https://github.com/casonk}{Cason Konzer}}}\XB \\
\end{center}
\hrulefill
\hfill
\newpage
```

```{r include=FALSE}
options(knitr.graphics.error = FALSE)

library(knitr)
library(ggplot2)
```

```{python include=FALSE}
import pandas as pd
import numpy as np
```


### 1. Find the distance between objects 1 and 3. Notice that we have mixed type of attributes. *(25 points)*

```{=latex}
\hrulefill

The following python code has will find the distance between any two objects for this dataset.
```

```{python}
# Build the dataframe.
table1 = {
  "Object Identifier": [1,2,3,4],
  "test-1 (nominal)" : ["A","B","C","A"],
  "test-2 (ordinal)" : ["excellent","fair","good","excellent"],
  "test-3 (numeric)" : [45,22,64,28]
  }
  
df1 = pd.DataFrame(table1)

# Set the index.
df1.set_index("Object Identifier",inplace=True)
df1
```

```{python}
# Replace nominal values.
df1["test-1 (nominal)"].replace("A",3,inplace=True)
df1["test-1 (nominal)"].replace("B",2,inplace=True)
df1["test-1 (nominal)"].replace("C",1,inplace=True)

t1_range = (df1["test-1 (nominal)"].max()-df1["test-1 (nominal)"].min())
df1["test-1 (nominal)"] = (df1["test-1 (nominal)"]-1)/t1_range
df1
```

```{python}
# Replace ordinal values.
df1["test-2 (ordinal)"].replace("excellent",3,inplace=True)
df1["test-2 (ordinal)"].replace("good",2,inplace=True)
df1["test-2 (ordinal)"].replace("fair",1,inplace=True)

t2_range = (df1["test-2 (ordinal)"].max()-df1["test-2 (ordinal)"].min())
df1["test-2 (ordinal)"] = (df1["test-2 (ordinal)"]-1)/t2_range
df1
```

```{python}
# Replace numeric values.
t3_range = (df1["test-3 (numeric)"].max()-df1["test-3 (numeric)"].min())
df1["test-3 (numeric)"] = (
  df1["test-3 (numeric)"]-df1["test-3 (numeric)"].min()
  )/t3_range
df1
```

```{python}
# Define our distance function.
def dist(i,j, df):
  neum = 0
  denom = 0
  for col in df.columns:
    print(col)
    if 'nominal' in col:
      if df[col].iloc[j-1] != df[col].iloc[i-1]:
        dist = 1
      else:
        dist = 0
    else:
      dist = np.abs(df[col].iloc[j-1] - df[col].iloc[i-1])
    print('feature distance:',dist)
    neum += dist
    denom += 1
    print()
  print('numerator:',neum)
  print('denomenator:',denom)
  print('\ntotal distance:')
  return neum/denom
  
# Compute the distance between object 1 & 3
dist(1,3, df1)
```

```{=latex}
We can see that the distance between object 1 and object 3 is \XBB $0.651$ \XB.

\newpage
```

### 2. Write a program in any language which can compute Manhattan and Euclidean distances between any two given vectors with any length. You can pass the length to your function, but please don’t limit the dimension to 2. You can test your function on vectors you fill in your code without asking user input. *(25 points)*

```{=latex}
\hrulefill

The following python code has independent functions that will find the Manhattan and Euclidean distance between any two given numpy vectors.
```

```{python}
# Create test vector 1.
test_v1 = np.array([0,1,2,3,5,4])
test_v1
```

```{python}
# Create test vector 2.
test_v2 = np.array([1,0,3,2,4,6])
test_v2
```

```{python}
# Build our Manhattan Function
def manhattaan(v1,v2):
  return np.sum(np.abs(v2-v1))

# Test our Manhattan Function
manhattaan(test_v1,test_v2)
```

```{python}
# Build our Euclidean function.
def euclidian(v1,v2):
  return np.sqrt(np.sum(np.square(v2-v1)))

# Test our Euclidean function.
euclidian(test_v1,test_v2)
```


```{=latex}
We can see that the simple test vectors pass for both functions.

\newpage
```

### 3. In the table below, determine whether passing a class has a dependency on attendance by using Chi-square test. *(25 points)*

```{=latex}
\hrulefill
```

```{python}
# Build the dataframe.
table2 = {
  ""     : ["Attended","Skipped","Total"],
  "Pass" : [25,8,33],
  "Fail" : [6,15,21]
  }
  
df2 = pd.DataFrame(table2)
df2

```

```{python}
# Total on attendance status.
df2["Total"] = df2["Pass"] + df2["Fail"]
df2

```

```{python}
# Create an expectation dataframe as a copy of the original.
expectation_df2 = df2.copy()
expectation_df2
```

```{python}
# Compute probabilities on attendance status 
expectation_df2["Total"] = expectation_df2["Total"].apply(
  lambda x: x / expectation_df2["Total"].max()
  )
expectation_df2
```

```{python}
# Compute passing expectations.
expectation_df2["Pass"] = (
  expectation_df2["Pass"].max()*expectation_df2["Total"])
expectation_df2
```

```{python}
# Compute failing expectations
expectation_df2["Fail"] = (
  expectation_df2["Fail"].max()*expectation_df2["Total"])
expectation_df2
```

```{python}
# Trim our original dataframe.
df2 = df2.iloc[:-1,:-1]
df2
```

```{python}
# Trim our expectation dataframe
expectation_df2 = expectation_df2.iloc[:-1,:-1]
expectation_df2
```

```{python}
# Compute passing portion of chi squared.
pass_ = np.sum(
  np.square(
    (df2["Pass"] - expectation_df2["Pass"])
    )/expectation_df2["Pass"]
  )
pass_
```

```{=latex}
\newpage
```

```{python}
# Compute failing portion of chi square.
fail_ = np.sum(
  np.square(
    (df2["Fail"] - expectation_df2["Fail"])
    )/expectation_df2["Fail"]
  )
fail_
```

```{python}
# Sum for our chi squared value.
chi_squared_ = pass_ + fail_
chi_squared_
```

```{=latex}
We can see that we yield \XBB $\chi^2 = 11.686$ \XB.

\newpage
```

### 4. In R, there is a built-in data frame called `mtcars`. Please calculate the correlation between `mpg` and `wt` attributes of `mtcars` by using `cor()` function. Then generate scatter plot based on these two attributes. *(25 points)*

```{=latex}
\hrulefill
```

```{r message=TRUE, warning=FALSE}
# Preview mtcars dataset.
mtcars
```

```{=latex}
\newpage
```

```{r}
# Set variables for naming.
mpg = mtcars$mpg
weight = mtcars$wt
```

```{r}
# Find Correlation.
cor(mpg, weight)
```

```{=latex}
We can see that the correlation between vehicle miles per gallon and vehicle weight in the mtcars dataset is \XBB $r_{mpg,weight} =  -0.868 $ \XB .
What this explains is that as one variable (weight/mpg) increases, the other (mpg/weight) decreases.
```

```{r}
# Make scatterplot.
plot(mpg, weight)
```

```{=latex}
\newpage
```

```{r}
# To coincide with the example given in the assignment. (:
plot(weight, mpg)
```